Quality of Service with Software Defined Network.

### 2017-03-11 report/paper.tex
Added content to paper

\usepackage[english]{babel}
\usepackage[T1]{fontenc}
\usepackage{cite,url,color} % Citation numbers being automatically sorted and properly "compressed/ranged".
\usepackage{graphics,amsfonts}
\usepackage{epstopdf}
\usepackage[pdftex]{graphicx}
\usepackage[cmex10]{amsmath}
% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally does.
\usepackage[utf8]{inputenc}
% Useful for displaying quotations
%\usepackage{csquotes}
% Compact lists
%\let\labelindent\relax
\usepackage{enumitem}

%tikz figures
\usepackage{tikz}
\usetikzlibrary{automata,positioning,chains,shapes,arrows}
\usepackage{pgfplots}
\usetikzlibrary{plotmarks}
\newlength\fheight
\newlength\fwidth
\pgfplotsset{compat=newest}
\pgfplotsset{plot coordinates/math parser=false}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Quality of Service with Software Defined Network.}

@@

Another limitation is the no default routing from a source to a destination: when an host
is plugged to a network of SDN switches, it won't know the destinatio. The only way it can
discover the host is by flooding the network with that packet, but this makes each switch raise a PacketIn message to the controller
slowing down the startup of the connection. This thing doesn't happen with traditional routing, because each router knows
the network disposition, but doesn't know each host how much traffic is doing. For that the administrator need to put a firewall.
The last one isn't needed, as the controller can set a switch to drop all the packets from/to a destination.

@@

\section{Needs and Goals}\label{sec:obb} %obbiettivi ed esigenze controllare
The main gool of this experience is to develop a small SDN network in a lab with real hardware
in orther to understand in a practical level how to implement a \textit{Quality of Service, \textbf{QoS}} service in a SDN scenarios.
Our network have to manage traffic with a global scope, in order to evolve in base on different parameters.
Moreover the SDN network need to interact with world wide web efficently.
%correggere
  \subsection{Adaptability of the network}
  Made Network Adaptable means poter dirottare il traffico in real time to ensure a lot of demands. For example we have to avoid link congestion,
  but we can prefer reroute some hosting traffic to worsen it's performance because  it is the traffic we don't want to encourage.
  To reach this gools we need to have a global knowledge on the network and we have to modify the routing table on every routerboard with a
  global scope.
  The performance of a single connection and entire network can ensure various type of metrics.
  In order to ensure the traffic shaping based on network informations, there is a need of a structure, 
  automatically updatable, where the link statistics can be saved to allow the best path for a data stream. The best metric will be described below.
  %  assicurare queste ci serve una struttura dati che si aggiorna costantemente nella quale immagazzinare le statistiche dei link
  % in modo all'occorrenza di trovare il percordo migliore per ogni singolo flusso dati.
  \#TODO MATTEO
  \subsection{Exterior comunication}
  \#TODO pitt
	Our network needed to be able to communicate with the external internet to be useful for a demo which demonstrate the capabilities of a SDN network.
	The first thing we needed to accomplish was to make the controller aware of what switch and which port of the switch is connected to the external internet:
	to do so, we allowed our controller to be able to receive in-line parameters. The two parameters we needed the controller to be aware were the mac address of the
	port of a switch that is connected to the link and the network address of our network, to let the controller know if a request from a host was meant to another host
	within the network or the data stream was intended to the internet.
	In this way the host connected to the inner network can surf the net regardless of which type of devices are switching and routing the traffic they are generating.

@@

\section{Results}\label{sec:results}
\subsection{Controller:}
We developed a controller that can handle a mesh network, which has loops, that avoid
the traffic flooding. The controller stores the switch and the connection between them;
knows which hosts are connected to each switch port, with the hypothesys that only one
host is connected at each port, unless the port is the one connected to the internet.
With the use of the statistics, the controller get the status of the network and, if necessary,
tries to adjust the traffic and put in the link with the higher path loss the torrent traffic,
make the connection between two gaming host the one with less delay and tries to redistribute all the traffic in a way
such that all the nodes have an average throughput as lowest as possible.
\\
\newline
We have used some principal function released with POX controller:
\begin{itemize}
 %\item \texttt{pox.topology.launch()} and \texttt{pox.openflow.topology.launch()} descrivere a che ci serve\cite{pox}.
 \item \texttt{pox.openflow.discovery.launch()} this function run algorithm in order to discovery the network and every time in witch some new 
 topological change occour it return an event with ( con cui ) we can update our topological knowledge.
 It does not provide the host event. For host event\cite{pox}.
 \item \texttt{pox.openflow.spanning_tree.launch()} this function is necessary with a presence of loop on topology 
 because with a spanning_tree algorithm it avoid the broadcast storm of message. When we set up \texttt{OFPP_FLOOD} like output port action in a table's 
 switch row if some spanning_tree algorithm run the \texttt{OFPP_FLOOD} is only theoretical infact the algorithm close some port for flooding service 
 on the switch in orther to avoid loop\cite{pox}.
\end{itemize}
\subsection{Real Lab Scenario}
The topological scenario in which we have set up experience is the next:
\begin{figure}[!h]
 \centering
 \includegraphics[scale=0.70]{images/topo.png}
 \caption{\emph{Experience scenario}}
 \label{fig:topo}
\end{figure}

@@

  \#TODO andrea
	For each statistic we got and each parameter we obtained, we changed the weight of the graphs. Every graph we used had a specific weight, so the
	Dijkstra algorithm for a weighted graph gave different minimum path between two hosts. From the parameter we saved, we choose to change the
	route to reach the destination and viceversa. The traffic follow the same path, in both the directions and depends only on the type of traffic:
	we choose to set the minimum unweighted path if no statistics were available, the minimum throughput path for the standard traffic, the minimum
	delay for the gaming one and the maximum packet error rate for the heavy traffic generated by a host.
	Two threshold were chosen based on the traffic analysis described on Sec. \ref{subsec:traf} which are 200 bytes/s for the gaming traffic and one megabyte/s
	when the node is considered in heavy traffic. Those two threshold were made as they are average values and considering that our network is made of few really fast link,
	so on one side the lower limit is sufficient not to consider gaming traffic to be a normal traffic, wherease it's easy to reach the high level of traffic load between two machines.
	The changing of the path were made only if the last change happened before one minute because the torrent traffic has, in its protocol, a discovery
	algortithm. This means that if a lot of changes were made, most of the traffic generated by the peer to peer, is the control one and the result is that the
	load in the network will be heavy and last longer.

@@

  During our analysis and in our tests, we prove that in a SDN mesh network, the ARP packets are useless.
  This type of packets is useful for a multipoint to multipoint connection where the connection is set between a
  HUB and a set of computers or between a bridge and a computer. The first case is no more in use in common networks as
  hiher the collision range. The bridges limits their collision range to the port. With openflow each switching unit can
  act as a bridge or as a router. In both cases analyses the header of the L2 packet L3 packet and some of the L4 header.
  If it has the rule to forward the packet, it will act as declared in the most specific matching rule. Otherwise it will
  send to the controller a PacketIn and the controller will act as programmed. If an host want to communicate to another host whithin
  the network, it sends the packet to the line and then the switch will check in its tables and execute the actions of the match found.
  For an IP network, it will check the path to reach a destination and then send to the defined output port. The mac address in this case
  became useless, as the connection is point to point. So, knowing the destination layer2 address is useless as the switch can act
  regardless of it. The only thing the switch need to know is the layer2, layer 3 addresses and, at most, the layer 4 address.
  This can be seen by a reader as a violation of the layers in the ISO/OSI stack, but to exploit the potentials of the SDN approach,
  this is necessary.

  \subsection{Default Gw problem}

  The computers don't need anymore to set a gateway to reach another network as the controller and the switches will redirect the traffic
  properly, according to the rules set by the first. The controller must only know which switch ports are connected to the
  internet and then add them to the path. in this way, the controller can set a path to reach it and provide connectivity with
  a traffic shaping policy and/or differentiate the traffic from and to different networks.
  \#TODO andrea

@@

  \#TODO andrea
	One of the problem encountered was that, if an host is already found by pox and there is a PacketIn event, the core of the controller
	switches the source and the destination IP and MAC address with the source one. The problem was partially solved, checking out the
	parameters known by the controller, i.e. if the two host are connected to two different switches and the controller switches the source with the
	destination address, we can switch it back; the result is that the PacketIn event now has correct informations.

>>>>

\usepackage[english]{babel}
\usepackage[T1]{fontenc}
\usepackage{cite,url,color} % Citation numbers being automatically sorted and properly "compressed/ranged".
\usepackage{graphics,amsfonts}
\usepackage{epstopdf}
\usepackage[pdftex]{graphicx}
\usepackage[cmex10]{amsmath}
% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally does.
\usepackage[utf8]{inputenc}
% Useful for displaying quotations
%\usepackage{csquotes}
% Compact lists
%\let\labelindent\relax
\usepackage{enumitem}
\usepackage{fancyhdr}
\fancyfoot[C]{\thepage}
%tikz figures
\usepackage{tikz}
\usetikzlibrary{automata,positioning,chains,shapes,arrows}
\usepackage{pgfplots}
\usetikzlibrary{plotmarks}
\newlength\fheight
\newlength\fwidth
\pgfplotsset{compat=newest}
\pgfplotsset{plot coordinates/math parser=false}

@@

\mainmatter
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Quality of Service with Software Defined Network.}

@@

Another limitation is the no default routing from a source to a destination: when an host
is plugged to a network of SDN switches, it won't know the destination. The only way it can
discover the host is by flooding the network with that packet, but this makes each switch raise a PacketIn message to the controller
slowing down the startup of the connection. This thing doesn't happen with traditional routing, because each router knows
the network disposition, but doesn't know each host how much traffic is doing. For that the administrator need to put a firewall.
The last one isn't needed, as the controller can set a switch to drop all the packets from/to a destination.

@@

\section{Needs and Goals}\label{sec:obb} %obbiettivi ed esigenze controllare
The main gool of this experience is to develop a small SDN network in a lab with real hardware
in orther to understand in a practical level how to implement a \textit{Quality of Service, \textbf{QoS}} service in a SDN scenarios.
Our network have to manage traffic with a global scope, in order to evolve in base on different parameters.
Moreover the SDN network need to interact with world wide web efficently.
\subsection{Connectivity}\label{sec:connectivity}
First thing to do is provide connectivity, letting the controller know the connection of a host and avoiding packet to be lost while the first
add the computer to its graphes and try to add the rules to let the traffic go from the source to the destination. This has been done doing a
\textit{controlled flooding}: when a PacketIn event arrives at the controller, this look into the data structure that saves the nodes connected.
\\ If source and destination nodes are already found, it set the flow rules from the source node' switch to the destination node switch. To the latter,
a PacketOut message is prepared with inside the packet arrived inside the PacketIn event and with destination the port of the destination switch connected
to the destination node. Then the message is sent to the destination switch, so that the first packet is not lost.
If, instead, the destination node is not found, the controller sends to the switch that generated the PacketIn event a PacketOut message
with destination all the ports of the switch itself and with payload, the message that arrived within the PacketIn message.
To avoid the network to be filled of the only packet sent, we kept the information regarding the switches on which the packet that was flooded:
if a PacketIn message arrives at the controller and both the packet and the switch got a PacketOut message within 2 seconds, the controller doesn't
do anything, as this is a loop.
\\ So this approach allowed us to make the network unaware that it's not a legacy one and allowed connectivity to be provided.
%correggere
\subsection{Adaptability of the network}
Made Network Adaptable means poter dirottare il traffico in real time to ensure a lot of demands. For example we have to avoid link congestion,
but we can prefer reroute some hosting traffic to worsen it's performance because  it is the traffic we don't want to encourage.
To reach this gools we need to have a global knowledge on the network and we have to modify the routing table on every routerboard with a
global scope.
The performance of a single connection and entire network can ensure various type of metrics.
In order to ensure the traffic shaping based on network informations, there is a need of a structure,
automatically updatable, where the link statistics can be saved to allow the best path for a data stream. The best metric will be described below.
%  assicurare queste ci serve una struttura dati che si aggiorna costantemente nella quale immagazzinare le statistiche dei link
% in modo all'occorrenza di trovare il percordo migliore per ogni singolo flusso dati.
\#TODO MATTEO
\subsection{Exterior comunication}
Our network needed to be able to communicate with the external internet to be useful for a demo which demonstrate the capabilities of a SDN network.
The first thing we needed to accomplish was to make the controller aware of what switch and which port of the switch is connected to the external internet:
to do so, we allowed our controller to be able to receive and process in-line parameters. We provided two parametes to the controller, so that
it could recognize where the traffic to the internet need to be sent. Theese two parameters were the mac address of the
port of a switch that is connected to the link and the network address of our network, to let the controller know if a request from a host was meant to another host
within the network or the data stream was intended to the internet. We provided only one gateway, as the Routeboard had few ports available for the packet switching
and only four SDN-capable switches to be able to simulate a real network that was generating traffic.

@@

\section{Results}\label{sec:results}
\subsection{Controller}
We developed a controller that can handle a mesh network, which has loops, that avoid
the traffic flooding. The controller stores the switch and the connection between them;
knows which hosts are connected to each switch port, with the hypothesys that only one
host is connected at each port, unless the port is the one connected to the internet.
With the use of the statistics, the controller get the status of the network and, if necessary,
tries to adjust the traffic and put in the link with the higher path loss the torrent traffic,
make the connection between two gaming host the one with less delay and tries to redistribute all the traffic in a way
such that all the nodes have an average throughput as lowest as possible.
\\
\newline
We have used some principal function released with POX controller:
\begin{itemize}
 %\item \texttt{pox.topology.launch()} and \texttt{pox.openflow.topology.launch()} descrivere a che ci serve\cite{pox}.
 \item \texttt{pox.openflow.discovery.launch()} this function run algorithm in order to discovery the network and every time in witch some new
 topological change occour it return an event with ( con cui ) we can update our topological knowledge.
 It does not provide the host event. For host event\cite{pox}.
 \item \texttt{pox.openflow.spanning_tree.launch()} this function is necessary with a presence of loop on topology
 because with a spanning_tree algorithm it avoid the broadcast storm of message. When we set up \texttt{OFPP_FLOOD} like output port action in a table's
 switch row if some spanning_tree algorithm run the \texttt{OFPP_FLOOD} is only theoretical infact the algorithm close some port for flooding service
 on the switch in orther to avoid loop\cite{pox}.
\end{itemize}
\subsection{Real Lab Scenario}
The topological scenario in which we have set up experience is the next:
\begin{figure}[!h]
 \centering
 \includegraphics[scale=0.70]{images/topo.png}
 \caption{\emph{Experience scenario}}
 \label{fig:topo}
\end{figure}

@@

\#TODO andrea
For each statistic we got and each parameter we obtained, we changed the weight of the graphs. Every graph we used had a specific weight, so the
Dijkstra algorithm for a weighted graph gave different minimum path between two hosts. From the parameter we saved, we choose to change the
route to reach the destination and viceversa. The traffic follow the same path in both the directions and depends only on the type of traffic:
we choose to set the minimum unweighted path if no statistics were available, the minimum throughput path for the standard traffic, the minimum
delay for the gaming one and the maximum packet error rate for the heavy traffic generated by a host.
Two threshold were chosen based on the traffic analysis described on Sec. \ref{subsec:traf} which are 200 bytes/s for the gaming traffic and one megabyte/s
when the node is considered in heavy traffic. Those two threshold were made as they are average values and considering that our network is made of few really fast link,
so on one side the lower limit is sufficient not to consider gaming traffic to be a normal traffic, wherease it's easy to reach the high level of traffic load between two machines.
The changing of the path were made only if the last change happened before one minute because the torrent traffic has, in its protocol, a discovery
algortithm. This means that if a lot of changes were made, most of the traffic generated by the peer to peer, is the control one and the result is that the
load in the network will be heavy and last longer.

@@

During our analysis and in our tests, we prove that in a SDN mesh network, the ARP packets are useless.
This type of packets is useful for a multipoint to multipoint connection where the connection is set between a
HUB and a set of computers or between a bridge and a computer. The first case is no more in use in common networks as
it higher the collision range. The bridges limits their collision range to the port. With openflow each switching unit can
act as a bridge or as a router. In both cases analyses the header of the L2 packet L3 packet and some of the L4 header.
If it has the rule to forward the packet, it will act as declared in the most specific matching rule. Otherwise it will
send to the controller a PacketIn and the controller will act as programmed. If an host want to communicate to another host whithin
the network, it sends the packet to the line and then the switch will check in its tables and execute the actions of the match found.
For an IP network, it will check the path to reach a destination and then send to the defined output port. The mac address in this case
became useless, as the connection is point to point. So, knowing the destination layer2 address is useless as the switch can act
regardless of it. The only thing the switch need to know is the layer2, layer 3 addresses and, at most, the layer 4 address.
This can be seen by a reader as a violation of the layers in the ISO/OSI stack, but to exploit the potentials of the SDN approach,
this is necessary.

\subsection{Default Gateway}
Analyzing the traffic while we were trying the controller on real machines in the lab, we have tried to set the default gateway to
other machines that were connected to the SDN network, but not to the internet. We proved that the controller can redirect correctly the
traffic to the internet. This means that
computers don't need anymore to set a gateway to reach another network as the controller and the switches will redirect the traffic
properly, according to the rules set by the first. The controller must only know which switch ports are connected to the
internet and then add them to the path. I this way, the controller can set a path to reach it and provide connectivity with
a traffic shaping policy and/or differentiate the traffic from and to different networks.

@@

During our tests, both with Mininet and in the lab, we found out that the core of POX, when a PacketIn event occoures, switches the source and the destination addresses,
both MAC and IP, if one of the hosts is already known. We weren't able to understand in which point of the code of POX  and why it happens.
% One of the problem encountered was that, if an host is already found by pox and there is a PacketIn event, the core of the controller
% switches the source and the destination IP and MAC address with the source one.
The problem was partially solved, checking out the
parameters known by the controller, i.e. if the two host are connected to two different switches and the controller switches the source with the
destination address, we can switch it back; the result is that the PacketIn event now has correct informations.
If the two host are in the same switch, the problem is solved using a controlled packet flooding, described in section \ref{sec:connectivity}
